\chapter*{Abstract}
\thispagestyle{plain}

Ensuring the anonymity of communication in diverse network systems is a technical challenge that requires solutions capable of achieving this objective while maintaining the usability of the network system. This thesis presents an analysis of anonymous communication networks (ACNs) in terms of their usability in various use case scenarios. The main goal of the work is to demonstrate that currently there is no universal ACN that is ideal for every use case scenario; each network performs differently depending on the specific requirements and the choice of ACN should be determined by the intended application. 

This work provides an overview of ACNs, focusing on their technical aspects and distinguishing the most prominent solutions today, as well as an overview of related work in terms of comparative analysis of ACNs. Furthermore, the work identifies potential use cases and application areas, categorises them based on the requirements that need to be fulfilled for a given use case, and presents potential threats for the ACNs.

To support the claim that different ACNs are optimal for different scenarios, a multi-criteria comparative analysis is performed, combining literature-based comparison with experimental measurements. This enables the evaluation of the most prominent ACNs today: Tor, I2P, Lokinet, and Nym, in terms of their usability and effectiveness across different use case groups.

The results of the comparative analysis clearly demonstrate that no single ACN is universally optimal; instead, the best choice depends on the specific use case. The work concludes with recommendations for the future development of ACNs.

To support a practical understanding of the findings of the comparative analysis, an educational demonstrator was developed. This demonstrator enables students to understand the results of the performed analysis through hands-on laboratory exercises and observe the strengths and weaknesses of various ACNs in terms of different application areas.

Keywords: Tor, I2P, Loki, Nym, anonymous communication network, ACN, analysis, use case, application area, usage
