\chapter{Threats, attacks and limitations}
\label{chap:Threats}

This chapter examines the threats faced by anonymous communication networks, focusing on passive and active attacks that compromise user anonymity and system integrity. It categorises and describes common attack vectors, evaluates how different ACNs respond to them, and highlights specific vulnerabilities to each network. The chapter also addresses broader limitations that are crucial for the long-term viability and accessibility of ACNs.

This chapter is important because understanding the range of threats is essential for evaluating existing ACNs and proposing any improvements. Only with a solid overview of real attack scenarios can we properly evaluate which ACNs are suitable for particular use cases and which are not. In addition, many of the threats presented here have already influenced the design of the most popular ACNs, with several countermeasures already in place. Recognising these defences is key before moving on to any comparison.

The first section discusses both passive and active attacks, describing in detail how they are carried out and how each ACN attempts to address them. The next part looks at economic sustainability as an often overlooked, but significant, limitation for ACNs going forward. Finally, the chapter examines the ongoing "censorship arms race" between ACN developers and those trying to block them. The presentation of these threats and countermeasures provides the necessary background for the comparative analysis in the following chapters and highlights the areas that require further development, which will be addressed in a dedicated section later on.

\section{Attacks}

Attacks on Anonymous Communication Networks (ACNs) can be broadly classified into two main categories: passive and active. \textbf{Passive attacks} on ACNs are mostly related to traffic analysis. These attacks are often extremely difficult to detect since the attacker does not interfere directly with the network traffic. \textbf{Active attacks}, on the other hand, involve some form of manipulation or interference with the network, such as injecting, modifying, or blocking messages. Active adversaries may attempt to compromise keys, introduce malicious nodes, perform denial-of-service attacks, or exploit protocol vulnerabilities. The following list contains several examples of the most prominent attacks on ACNs for both categories:

\begin{itemize}
    \item Passive
    \begin{itemize}
        \item Timing correlation
        \item Size/volume correlation
        \item Intersection attack/statistical disclosure/traffic pattern observation
        \item Browser fingerprinting
        \item Unencrypted content observation
    \end{itemize}
    \item Active
    \begin{itemize}
        \item Key exposure
        \item Tagging attack
        \item Content modification
        \item Sybil attack
        \item N - 1 attack
        \item Denial of Service (DoS)
        \item Compulsion attack
        \item Cryptographic attacks
        \item Vulnerabilities exploit
        \item ACN-specific attacks
    \end{itemize}
\end{itemize}

\textbf{Timing correlation} is one of the most effective attacks against low latency ACNs. Given an adversary who can watch traffic entering and leaving the network, he can easily distinguish who sent the message due to the timing correlation. In order to mitigate this kind of attack, a non-deterministic strategy needs to be included. In Mix-nets, batching is implemented with various strategies. Addressing the timing correlation is always associated with additional latency and, therefore, smaller usability in lower-latency usage scenarios. While there are propositions of optional inclusion of non-trivial delays and batching strategies in I2P, they are not yet implemented. As in I2P, each user acts as a router, it will impose additional effort on the attacker to distinguish the traffic originating from the user from the traffic he is passing through; however, it is not a complete mitigation of this type of attack. Tor and Lokinet are the most vulnerable to the timing correlation attack; however, the more users the network has, and therefore the larger anonymity set is, the harder it is to perform this attack.

\textbf{Size correlation} is an attack in which an adversary correlates the message with its sender due to its size. Lokinet is the most vulnerable to this type of attack as it does not propose any mitigation. Tor uses fixed-size messages; however, it is only a partial mitigation of the issue, especially when a user is sending large messages, as the sequential chunks clearly form a larger message, and the pattern can be, therefore, observed. In I2P, the chunks can be initially combined into larger messages and then, after leaving the sender’s outbound tunnel, split and delivered through the inbound tunnels of multiple receivers, which can make the size correlation harder to perform. Nym provides the best defence against this attack, providing both padded Sphinx messages along with cover traffic, which makes observing message size patterns impossible.

Identifying users by correlating on-line presence over time is known as \textbf{intersection attack/statistical disclosure/traffic pattern observation}. When an adversary can observe one endpoint of the communication as well as the fact of a user entering the network, the fact of intensified message number on the receiving side can be associated with a certain user that is most likely a sender. This attack is especially feasible when an adversary can observe the network for a long period of time. In theory, when a user is constantly connected to the ACN, it would not be distinguishable; however, this is not an acceptable assumption to make. The more users and routers are on the network, the more complex the attack is to be performed. Masking traffic and non-public routers, for example pluggable transports and bridges from Tor network, can also mask a fact of connecting to the network. The best solution to this issue, deployed by Nym, is to include a cover traffic - as every endpoint is constantly receiving it, an adversary is not able to tell whether a certain entity is receiving larger traffic than usual. Batching and mixing are another potential solution to this problem.

\textbf{Browser fingerprinting} is a browser-related attack that involves the collection of user data that can later be used to identify him. Tor solves this issue by providing a custom, hardened browser. Although other ACNs are vulnerable to this issue, it is less performable in hidden service-focused ACNs like I2P.

\textbf{Unencrypted content observation} is an attack associated with malicious exit nodes that can observe unencrypted traffic from users that reach clearnet resources over unsafe protocols such as HTTP. The solution for this attack is either to only use end-to-end encrypted protocols, for example, HTTPS-based websites, or not to leave the ACN at all and focus on the hidden services.

Although brute-force-based attacks on cryptographic keys are not feasible in an acceptable time with current technology, the key can be acquired by other means. In case of keys that are not periodically rotated and therefore ACNs that do not provide perfect forward secrecy, it may be potentially used to decrypt an encryption layer from the past. This attack is known as \textbf{key exposure}. In most of the ACNs, the keys are ephemeral and therefore leaked keys can only be used to unwrap one layer of encryption for a short period of time.

\textbf{Tagging attack} is an attack based on modifying a message so that it would be distinguishable later in the message path. Message integrity checking in ACNs is crucial when it comes to defending against this type of attack, and most of them include the mechanisms for providing integrity.

\textbf{Content modification} is an attack that is one step further compared to the passive unencrypted content observation attack described above. In this case, the malicious exit node modifies the content before passing it through. The mitigation is the same as described with the unencrypted content observation attack.

\textbf{Sybil attack} is an attack in which an adversary creates a large number of routers and tries to, for example, deanonymise users or perform other attacks such as DoS. I2P is particularly vulnerable to this attack due to its peer-to-peer architecture and the lack of effective mitigation mechanisms. In Tor, the attack needs to be performed over a long period of time as a sudden spike in the number of nodes can be easily noticed and addressed. Lokinet and Nym consider themselves to be more resistant to this type of attack due to economic incentives for the node runners. Nym also includes penalisation of the operators that register multiple nodes.

\textbf{N - 1 attack} is an attack aimed at Mix-nets in which an adversary floods the mix with his n - 1 messages, leaving space for one non-adversary message which he aims to target. Nym addresses this issue by using cover traffic and stratified topology.

\textbf{Partitioning attack and eclipse attack} are attacks in which an adversary tries to isolate a single target (eclipse attack) or a larger network segment (partitioning attack), forcing victims to connect to the routers controlled by the adversary. The attack involves interference with the mechanisms for discovering routers within the network, for example, netDb from I2P or Tor directory servers. ACNs such as I2P with more decentralised nature during new user joining, can also suffer from a specific type of the eclipse attack called the bootstrap attack in which an attacker provides a list of his malicious routers instead of the legitimate ones.

Attack that aims to disrupt the system, for example an ACN, is known as \textbf{Denial of Service (DoS)}. The attack has its distributed version - DDoS, which is performed from many places. The larger the network, the more costly the attack becomes. In peer-to-peer ACNs like I2P naive DDoS may potentially have the opposite effect, as each new user will also act as a router and speed-up the network, unless the attacker modifies his nodes to only consume from the network and do not contribute. There are many forms of DoS, often addressing ACN-specific technical solutions, for example, the Sniper attack \cite{sniper} on Tor.

\textbf{Compulsion attack} is a non-technical type of attack in which a router runner is forced to give up control over his router in order for the adversary to control the traffic passing through it. This attack is not useful against networks that have perfect forward secrecy, which means frequent key rotation, as gaining control over nodes is not useful for decrypting previous messages. Although this attack may have been more dangerous if the node runner ran all nodes in a given path, most of the ACNs provide diversification mechanisms for the nodes in order to avoid them from being under the same jurisdiction et cetera.

While today's ACNs use algorithms considered safe, meaning that they cannot be brute forced in an acceptable time, the highest danger is associated with quantum computers and their ability to break currently safe cryptographic algorithms; therefore performing \textbf{cryptographic attacks}. While it is not a threat that is present today, Tor and I2P are already in the process of replacing their currently used algorithms with post-quantum-based ones.

There is also a possibility of \textbf{code vulnerabilities exploits}, however such attacks are mostly related to less mature ACNs with less research backup.

There are many \textbf{ACN-specific attacks} that are usually an effect of omitting a certain aspect during the projection of the network. Contrary to bugs, in this example, the ACN works as design but the issue lies with the design itself. The solution is also more costly than a resolution of a bug as it involves changes in the architecture. An example can be the guard discovery attack \cite{guard-discovery} on Tor.

\section{Economic sustainability}
Economic sustainability is a critical challenge in ACNs. The approach to the economic issues related to ACNs varies from network to network. Peer-to-peer solutions, such as I2P, benefit from the decentralised nature. Each I2P user who uses the network contributes to it for the others, enhancing scalability and anonymity. The I2P development team does not accept direct donations; they encourage supporting application developers. One of the examples is StormyCloud, which provides a free outproxy service for I2P.

Tor has an entirely altruistic volunteer-based model in terms of running nodes. This approach has turned out to be working well for more than two decades of Tor history. Tor underlines the importance of diversity among their nodes, and the approach they have chosen seems to align with this goal.

Lokinet focuses on economic incentives for node runners, rather than volunteer-based networks. Although this approach can attract a broad set of participants, as it currently does, it introduces several potential challenges. Firstly, node runners can choose the cheapest infrastructure providers in order to maximise their profit, which can lead to less diversity of nodes within the network. Secondly, the token-based reward that Lokinet proposes may be volatile. The value of the token they provide is strictly speculative and may vary significantly over time, potentially discouraging node runners when the price goes down. The development team does not accept direct donations; similarly, for node runners, they benefit from the tokens.

The Nym network has a similar approach to Lokinet in terms of incentives for node runners. Although this can encourage more node runners, it shares the same risks, including token volatility and centralisation issues. Unlike other ACNs, Nym requires a monthly fee for network usage that funds the development behind Nym.

\section{Censorship arms race}
Efforts to circumvent Internet censorship using ACNs are continually met with countermoves by restrictive governments, creating an ongoing technological arms race. Tor is in particular an object of such actions \cite{tor-arms-race}, as it is the most popular ACN today. Governments try to block ACNs by targeting their management mechanisms, public relay nodes for client-server ACNs, websites from which you can download a given ACN, or by using deep packet inspection (DPI) to detect ACN-specific traffic. In response, the Tor Project has developed solutions such as bridge relays and pluggable transports to help users access the network even under heavy censorship. Although censors have access to significant resources and increasingly sophisticated techniques, the Tor Project adapts by introducing new technical measures, making censorship circumvention a continuous and evolving challenge. Other ACNs have not yet deployed such advanced censorship circumvention techniques; although it is often a part of their roadmap.

\section{Summary}
This chapter has provided an overview of the key threats, attacks, and limitations associated with anonymous communication networks. Passive and active attacks were discussed, along with countermeasures that are implemented in current ACNs. Among these, the threat of traffic analysis remains significant, as it cannot be fully mitigated by currently popular onion routing-based solutions, while Mix-net-based approaches offer better protection against and therefore increased anonymity, as traffic analysis and the attacks based on it can significantly hurt the provided level of anonymity. Other threats were also considered, not only in terms of passive and active attacks but also about economic incentives and censorship. The issues mentioned within this chapter set the practical boundaries of ACN security and will be crucial for the comparative analysis in the next chapter. Moreover, addressing the open problems and limitations identified in this chapter will be essential for the future development of ACNs, and these will be discussed in detail in a dedicated section later in the thesis.
